The Unicode Cookbook \linebreak for Linguists

### 2018-06-05 book/chapters/implementation.tex
fixes abbreviations, address paperhivers through chapter 1

These executables can be 
used in the resulting file path, or they can be linked and/or copied to any location as wanted. 
For example, a good way to use the executables in a terminal is to
make softlinks (using \texttt{ln}) from the executables to a directory in your
PATH, e.g. to \texttt{/usr/local/bin/}. The two executables are named
\texttt{tokenize} and \texttt{writeprofile}, and the links can be made directly 
by using Rscript to get the paths to the executables within the terminal.

@@

Some characters look identical, although they are encoded differently.
Unicode offers different ways of normalization (see
Section~\ref{pitfall-canonical-equivalence}), which can be invoked here as well
using the option \texttt{normalize}. NFC normalization turns everything into the
precomposed characters, while NFD normalization separates everything into base
characters with combining diacritics. Splitting by code points (i.e. \texttt{sep~=~""}) 
shows the results of these two normalizations in Tables \ref{tab:profile3} \& \ref{tab:profile4}.

@@

Everything is not yet correct with the graphemic parsing of the example discussed
previously. The sequence <sh> in `mishap' should not be a digraph, and
conversely the sequence <sch> in `mischief' should of course be separated into
<s> and <ch>. One of the important issues to get the graphemic parsing right is
the order in which graphemes are parsed. For example, currently the grapheme
<sch> is parsed before the grapheme <ch>, leading to <m\ i\ sch\ ie\ f> instead
of the intended <m\ i\ s\ ch\ ie\ f>. The reason that <sch> is parsed before
<ch> is that by default longer graphemes are parsed before shorter ones. Our
experience is that in most cases this is expected behavior. You can change the
ordering by specifying the option \texttt{ordering}. Setting this option to
\texttt{NULL} results in no preferential ordering, i.e. the graphemes are parsed
in the order of the profile, from top to bottom. Now `mischief' is parsed
correctly, but `scheme' is wrong. So this ordering is not the solution in this
case.

@@

By specifying more than one ordering, these orderings are used to break ties,
e.g. the default setting \texttt{ordering = c("size", "context", "reverse")}
will first order by size, and for those with the same size, it will order by
whether there is any context specified or not. For lines that are still tied
(i.e. have the same size and both/neither have context) the order will be
reversed compared to the order as attested in the profile, because most
hand-written specifications of graphemes will first write the general rule,
followed by more specific regularities. To get the right tokenization, these 
rules should in most cases be applied in reverse order.

@@

In all examples above we have used the default formatting for grapheme
separation using space as a separator, which is obtained by the default setting
\texttt{sep~=~"~"}. It is possible to specify any other separator here,
including the empty string, i.e. \texttt{sep = ""}. This will not show the
graphemic tokenization anymore (although it has of course been used in the
background).

>>>>

These executables can be 
used in the resulting file path, or they can be linked and/or copied to any location as wanted. 
For example, a good way to use the executables in a terminal is to
make softlinks (using \texttt{ln}) from the executables to a directory in your
PATH, e.g.\ to \texttt{/usr/local/bin/}. The two executables are named
\texttt{tokenize} and \texttt{writeprofile}, and the links can be made directly 
by using Rscript to get the paths to the executables within the terminal.

@@

Some characters look identical, although they are encoded differently.
Unicode offers different ways of normalization (see
Section~\ref{pitfall-canonical-equivalence}), which can be invoked here as well
using the option \texttt{normalize}. NFC normalization turns everything into the
precomposed characters, while NFD normalization separates everything into base
characters with combining diacritics. Splitting by code points (i.e.\ \texttt{sep~=~""}) 
shows the results of these two normalizations in Tables \ref{tab:profile3} \& \ref{tab:profile4}.

@@

Everything is not yet correct with the graphemic parsing of the example discussed
previously. The sequence <sh> in `mishap' should not be a digraph, and
conversely the sequence <sch> in `mischief' should of course be separated into
<s> and <ch>. One of the important issues to get the graphemic parsing right is
the order in which graphemes are parsed. For example, currently the grapheme
<sch> is parsed before the grapheme <ch>, leading to <m\ i\ sch\ ie\ f> instead
of the intended <m\ i\ s\ ch\ ie\ f>. The reason that <sch> is parsed before
<ch> is that by default longer graphemes are parsed before shorter ones. Our
experience is that in most cases this is expected behavior. You can change the
ordering by specifying the option \texttt{ordering}. Setting this option to
\texttt{NULL} results in no preferential ordering, i.e.\ the graphemes are parsed
in the order of the profile, from top to bottom. Now `mischief' is parsed
correctly, but `scheme' is wrong. So this ordering is not the solution in this
case.

@@

By specifying more than one ordering, these orderings are used to break ties,
e.g.\ the default setting \texttt{ordering = c("size", "context", "reverse")}
will first order by size, and for those with the same size, it will order by
whether there is any context specified or not. For lines that are still tied
(i.e.\ have the same size and both/neither have context) the order will be
reversed compared to the order as attested in the profile, because most
hand-written specifications of graphemes will first write the general rule,
followed by more specific regularities. To get the right tokenization, these 
rules should in most cases be applied in reverse order.

@@

In all examples above we have used the default formatting for grapheme
separation using space as a separator, which is obtained by the default setting
\texttt{sep~=~"~"}. It is possible to specify any other separator here,
including the empty string, i.e.\ \texttt{sep = ""}. This will not show the
graphemic tokenization anymore (although it has of course been used in the
background).
